Quantitative Comparison With Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions

### 2023-05-02 MANUSCRIPT/manuscript.tex

\documentclass[journal=jctcce]{achemso}
\usepackage[version=3]{mhchem} 
\usepackage[T1]{fontenc}       
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{xr}
\usepackage{multirow}
% \usepackage{rotating} 
% \usepackage{times}
% \usepackage{graphicx}
% \usepackage{setspace}
% \usepackage{amsmath}
% \usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}

@@

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{matti.javanainen@helsinki.fi}
%
\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\alsoaffiliation{University of South Florida}
%
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{MPI}
%
\author{O. H. Samuli Ollila}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{samuli.ollila@helsinki.fi}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
%
\author{Thomas Piggot}
\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}

@@

\todo{Update Author list \& Affiliations}

\begin{abstract}
Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, and thereby renders the membrane stiffer. Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models predict quantitatively different behavior. Such limitations are easily neglected, as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions. Still, for their validity, it is crucial that the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol, are modelled accurately.
%
Here, we perform a systematic comparison of the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol. Comparison of lipid bilayer simulations with deuterium order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations. The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively. 
%The performance of the commonly used CHARMM36 model is subpar, as it overestimates diffusion at low cholesterol concentrations, and gets excessively ordered upon the addition of cholesterol. 
The accompanied X-ray scattering data set measured for this work will foster future force field development and refinement.
\end{abstract}

@@

\todo{Check that the uses of ``model'' and ``force field'' are consistent}

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation}, %\textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, 
in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.

While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations are widely used to provide a detailed view of complex membranes and the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}. This is further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable the simulations of ever more complex and realistic membranes .

The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively. Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines that enables rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.

While simulating of complex membranes with cholesterol have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge.
%relies on the quality of the used model. 
%Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons. Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature. This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.
%
Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C-H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}. 
%This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. 
This approach has been useful for finding the best models for headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE) \cite{bacle2021inverse} and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}. Furthermore, quantitative quality measures based on C-H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}. However, such automatic quality evaluation is limited to simulations that have experimental data available with corresponding composition and temperature. 
%One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and 
Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automized in the NMRlipids databank.

Here, we demonstrate how simulations of binary POPC-cholesterol mixtures can be evaluated against experimental NMR and X-ray scattering data
%order parameters and electron density profiles can be readily 
by interpolating through multiple CHOL concentrations. Because effect of cholesterol on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously~\cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in cholesterol induced lateral membrane heterogeneity.
%to intermediate CHOL concentrations. This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison. 
We also evaluated 
%whether the simulation force fields reproduce 
the dependence of lateral diffusion coefficients on cholesterol against pulsed field gradient (PFG) NMR experiments~\cite{filippov2003effect,filippov2003influence} using the recent theoretical work that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of popular force fields at different CHOL concentrations. We then used the best-performing force fields to analyze molecular organization at different cholesterol concentrations, as we expect them to best reflect this organization in reality.
%
%Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby 
While we focus here on POPC-cholesterol mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. 
%Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime. 
%Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}. In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer. The form factor and electron density profiles can be compared with the same profiles calculated from the simulation data. Whereas NMR provides data on the conformations of individual lipids, form factors from scattering captures the details of the structure of the entire lipid bilayer \cite{ollila16}.

@@

\subsection{X-ray Scattering Experiments}
Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high} . This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content. Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification. All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria). The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtained both contributions. The electron density profiles were modelled from form factors using the SDP model where volume distribution functions are modelled by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}. 
%CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. 
Cholesterol was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
%Additional figures show the volume distribution functions and the resulting electron density profiles.
Membrane thickness was defined as twice the distance from the electron density maximum to the membrane center. The electron density maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.

@@

We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%. Systems were simulated using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256 or 1024 POPC molecules). 
%The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules). 
The number of POPC molecules was kept constant across the different CHOL concentrations. All membranes were solvated by 50 waters per lipid (POPC or CHOL).  All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog). All simulations were stored to the NMRlipids databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}.

\todo{Run Lipid21 for revision?}

@@

\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles 
%were calculated using the \texttt{form\_factor.py} script in 
automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank} 
%(\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts. 
were used. Similarly to the experimental X-ray scattering data, membrane thickness was defined as twice the distance from the electron density maximum to the membrane center. Locations of maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet. 
%
%\paragraph{Order Parameters:} The C--H order parameters were calculated as
%%
%\begin{align}
%    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
%\end{align}
%%
%where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane. 
To simplify the interpolation, C-H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged. These groups rotate freely and thus the order parameters are essentially identical for both (all) hydrogens in experiments and simulations.
%, the values are also very similar, justifying this approach. 
An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses. The C--H order parameters for the POPC head group 
%were calculated similarly, but values for both hydrogens 
are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.

\paragraph{Lateral Diffusion Coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet with the \texttt{gmx msd} tool. The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
%
\begin{align}
	\mathrm{MSD}=4D_\mathrm{PBC}\Delta.
\end{align}

The diffusion coefficients extracted from the three simulation box sizes were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $H$ is half the thickness of the water layer, $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length, and $\mu_\mathrm{m}$ and $\mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively~\cite{vogele2018hydrodynamics}. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%). The final simulation box dimensions were used for $L$ and $H$. For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.

>>>>

\documentclass[journal=jctcce]{achemso}
\usepackage[version=3]{mhchem} 
\usepackage[T1]{fontenc}       
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{xr}
\usepackage{multirow}

@@

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{matti.javanainen@helsinki.fi}
%
\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\alsoaffiliation{University of South Florida}
%
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{Department of Chemistry, University of Bergen, Norway}
\alsoaffiliation{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}

%
\author{O. H. Samuli Ollila}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{samuli.ollila@helsinki.fi}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
%
\author{Thomas Piggot}
\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

%\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}
\title{Quantitative Comparison With Experiments Reveals Imperfections in Force Fields' Descriptions of Phospholipid--Cholesterol Interactions}

@@

\begin{abstract}
Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, renders the membrane stiffer, and drives domain formation. Molecular dynamics simulations have played a crucial role in resolving these effects at molecular level, yet it has recently become evident that different simulation force fields predict quantitatively different behavior. Although easily neglected, identifying such limitations are becoming increasingly important as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions where it is crucial to accurately model the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol.
%
Here, we define quantitative quality measures for binary lipid mixtures in membranes against C--H bond order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering. We then perform a systematic evaluation on how well the commonly used force fields describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol. None of the tested force fields clearly outperforms the others across the tested properties and conditions, yet Slipids parameters provide the best overall performance.

The quality evaluation metrics introduced in this work will particularly foster the future force field development and refinement for multi-component membranes with automated approaches.
\end{abstract}

@@

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between other lipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation}, in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane,manna2016mechanism}.

While the structure and dynamics of heterogeneous membranes are difficult to capture experimentally, atomistic resolution molecular dynamics (MD) simulations have been employed to obtain a detailed view on the lateral organization of lipid bilayers driven by lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed,enkavi2019multiscale,marrink2019computational}. Such efforts are further facilitated by an increasing selection of available force fields with compatible lipid and protein parameters that enable the simulations of ever more complex and thus realistic membranes \cite{lorent2020plasma}.

The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the force field by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig16}, respectively. Notably, simulations using CHARMM36, Lipid17, and Slipids can now be set up using CHARMM-GUI for multiple simulation engines, which enables the rapid set up of complex membrane simulations \cite{lee16,lee2020charmm}.

While simulations of complex membranes with CHOL have become relatively straightforward task, estimating the trustworthiness of MD simulations for complex membranes remains a challenge. Our earlier work has demonstrated that the conformational ensembles of lipids from an MD simulation can be evaluated against C--H bond order parameters from solid state NMR experiments \cite{botan15,ollila16,catte2016molecular,antila2019headgroup,bacle2021inverse}. This approach has been useful for finding the best force fields for the description of the headgroups of phosphatidylcholine (PC)~\cite{botan15}, phosphatidylserine (PS)~\cite{antila2019headgroup}, phoshatidylethanolamine (PE)~\cite{bacle2021inverse} and phosphatidylglycerol (PG)~lipids \cite{bacle2021inverse}, for evaluating and improving membrane interactions with ions~\cite{catte2016molecular,antila2019headgroup,bacle2021inverse,melcr18,melcr19} and small molecules~\cite{nencini22}, and for finding simulation parameters predicting most realistic packing properties of membranes~\cite{antila2022emerging,NMRlipidsDatabank}. Furthermore, quantitative quality measures based on C--H bond order parameters have been recently defined and used to rank simulations in the NMRlipids databank~\cite{NMRlipidsDatabank}. However, such automatic quality evaluation is limited to simulations for which experimental data at a corresponding composition and temperature are available. Because simulations mimicking all experimental compositions for multi-component membranes are often tedious to produce, quality evaluation of mixed lipid bilayers is not yet fully automatized in the NMRlipids databank.

Here, we demonstrate how simulations of binary POPC--CHOL mixtures can be evaluated against experimental NMR and X-ray scattering data by interpolating through multiple CHOL concentrations. Because effect of CHOL on lipid headgroup and its decoupling of acyl chain ordering has been discussed previously \cite{botan15,antila22b}, we focus here on acyl chains that are expected to play larger role in CHOL induced lateral membrane heterogeneity. We also evaluated the dependence of lateral diffusion coefficients on CHOL against pulsed field gradient (PFG) NM experiments \cite{filippov2003effect,filippov2003influence} using the recent theoretical framework that allows a quantitative comparison with experiment after eliminating the finite-size effects in MD simulations~\cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of popular force fields at different CHOL concentrations. While we focus here on POPC-CHOL mixture, we expect our results to set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems.

@@

\subsection{X-ray Scattering Experiments}

Fully hydrated multilamellar vesicles (MLVs), composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments, were prepared for small angle X-ray scattering (SAXS) experiments using rapid solvent exchange as described previously \cite{rieder2015optimizing,belivcka2017high}. This avoids the precipitation of CHOL crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% CHOL content. Lipids, purchased from Avanti Polar Lipids (Alabaster, AL, USA), were used as dry powders without any further purification. All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria). The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtained both contributions. The electron density profiles were modelled from form factors using the SDP model where volume distribution functions are modelled by individual Gaussians or error functions~\cite{heberle12,Kucerka08a,kucerka12}. CHOL was described using two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. Membrane thickness was defined as twice the distance from the electron density maximum to the membrane center. The electron density maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data.

@@

We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%. Systems were simulated using four commonly used force fields, namely CHARMM36 (often abbreviated ``C36'' in figure legends in this work) \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig16} force fields. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes (64, 256 or 1024 POPC molecules). The number of POPC molecules was kept constant across the different CHOL concentrations. All membranes were solvated by 50 waters per lipid (POPC or CHOL). The small membranes (with 64 lipids) were first generated using CHARMM-GUI and equilibrated using the standard protocols for CHARMM36, Slipids, and Lipid17 for which inputs are readily available from CHARMM-GUI \cite{lee16,lee2020charmm}. Then, the atomic coordinates were replicated in the membrane plane to create the 4- and 16-fold larger simulation systems. Since CHARMM-GUI does not support MacRog, the production simulations were initiated from equilibrated CHARMM36 structures, since they share the atom ordering. All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The first 10~ns of each simulation were omitted from analyses. The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog). All simulations were stored to the NMRlipids databank~\cite{NMRlipidsDatabank} with the ID numbers listed in Table~\ref{IDtable}. The CHARMM36 simulations have been previously analyzed for their dynamic properties in Ref.~\citenum{fabian2023protein}.

@@

\paragraph{Structural properties:} The C--H bond order parameters, form factors and electron density profiles, automatically calculated by the NMRlipids Databank~\cite{NMRlipidsDatabank}, were used. Similarly to the experimental X-ray scattering data, membrane thickness was defined as twice the distance from the electron density maximum to the membrane center. Locations of maxima were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.

To simplify the interpolation of order parameter data to intermediate CHOL concentrations (see \textit{``Quantitative quality evaluation of the effect of CHOL on membrane properties''} below), C--H bond order parameters of 2 (3) hydrogens in CH$_2$ (CH$_3$) groups in the POPC acyl chains were averaged. These groups rotate freely and thus the order parameters are essentially identical for both (all) hydrogens in experiments and simulations. An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both simulations and experiments, and were thus treated separately in our analyses. The C--H order parameters for the POPC head group are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.

\paragraph{Lateral Diffusion Coefficients:} Lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet with the \texttt{gmx msd} tool. The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
%
\begin{align}
	\mathrm{MSD}=4D_\mathrm{PBC}\Delta.
\end{align}

The diffusion coefficients extracted from the three simulation box sizes (and thus membrane edge lengths $L$) were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}$ is the Boltzmann constant, $T$ is the temperature, $H$ is half the thickness of the water layer, $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffman--Delbr\"{u}ck length, and $\mu_\mathrm{m}$ and $\mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively~\cite{vogele2018hydrodynamics}. Thus, the fit of Eq.~\eqref{eq:sdlatpbc} to the $D_\mathrm{PBC}$ values calculated from the simulation as a function of simulation box size $L$ has two free parameters, namely $D_\infty$ and $\mu_\mathrm{m}$, both of which are free of finite size effects and can thus be compared to experiment. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of $\mu_\mathrm{f}= 0.3228$~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%). The simulation box dimensions, membrane edge length $L$ and the dimension normal to the membrane plane ($z$) $L_z$, were taken from the final configuration of each simulation. $L_z$ was only needed in the calculation of $H$ (see below). Membrane thickness $h$ was obtained as described in \textit{``Structural properties''} above. As $h$ and $H=(L_z-h)/2$ are constants in Eq.~\eqref{eq:sdlatpbc}, the average values from the three systems sizes were used in the fits.

\paragraph{Quantitative quality evaluation of the effect of CHOL on membrane properties:} To ease the evaluation of simulations against experimental data with non-matching CHOL concentrations, we interpolated the effect of CHOL in simulations and experiments for CHOL concentrations ranging from 0\% to 46\%. 2D matrices were created by interpolation for the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains and CHOL concentration) using the \texttt{interp2} function in Matlab based on linear interpolation. Similar matrices were also generated for the electron density profiles (as a function of normal distance from bilayer center and CHOL concentration) but with the \texttt{scatteredInterpolant} function in Matlab based on Delanuay triangulation \cite{amidror2002scattered}. Linear 1D interpolation with the \texttt{interp1} function in Matlab was used for the CHOL-dependence of the 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations and diffusion coefficients. These interpolations were then used to calculate deviations (in \%) from experimental values across CHOL concentrations to quantify the quality of the lipid force fields.  For the 2D matrices, the absolute values of the differences between matrices from simulations and experiments were first calculated. The averages of differences over the carbon atoms in the acyl chain (order parameters) or across the entire $z$ coordinate in the simulation (density profiles) were then calculated. This resulted in 1D deviation vectors as a function of CHOL concentration. For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration. All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \% between a simulation and experiments as a function of CHOL concentration. For C--H bond order parameters, also the deviation matrices between simulations and experiments were used to illustrate quality of simulations.
